Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-28 numerical.tex
updating due to barker latter 2015

\subsubsection{Comment on `surface modes'}\label{surf_comment} 
Our fiducial setup with $\khat=10$ in 
Fig. \ref{compare_modes_iso_kx10} (black dots)  lack `surface modes'
found by \cite{nelson13}, \cite{mcnally14}  and 
by \cite{barker15}.  We illustrate this with the mode
diagram for $\khat=30$ in Fig. \ref{compare_modes_iso_kx30}. Surface
modes lie almost parallel to the vertical axis
in such diagrams. They are readily identified for  $\zmax\geq 5H$ and
are marked by `s'.  We typically find an increasing
number of them with increasing $\zmax$ and/or $\khat$,
consistent with \cite{barker15}.

As explained by \cite{barker15}, the occurence of surface modes in the
vertically isothermal disk is due to imposing a finite vertical domain
size. The fastest growing mode is typically a surface mode, if they
exist, but their association with vertical boundaries means that
$\mathrm{max}(\nu)$ depends on the choice of $\zmax$. Thus, the
vertically isothermal disk is not useful for studying surface modes. 
%Thus, seeking the most unstable mode in a vertically isothermal disk
%may not yield robust results because this

@@

We find the number of unstable modes decrease with increasing
$\beta$. Consider the fiducial setup with $\zmax=5H$ 
(middle panel). The fundamental mode ($M=0$) is not the most unstable for   
$\beta=0.01$ as there are modes with $M>0$ that have larger growth 
rates. However, by comparing with results for $\zmax=3H$, we see these
higher order modes are strongly influenced by boundary conditions, and
there is a significant mismatch between numerical and analytical
results for $\zmax=3H$ for $M>1$.

However, the fundamental mode is dominant for $\beta \geq 
0.03$. Higher order modes, with larger growth rates at small $\beta$,
are more effectively stabilized with increasing $\beta$. For
$\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we only find the
fundamental mode. This is consistent with the discussion in 
\S\ref{iso_vsi_beta_crit}.

@@

Fig. \ref{lowfreq_eigenfunc_cool}---\ref{lowfreq_eigenfunc_2d_cool}
shows the fundamental mode with $\khat=10$ in a disk with 
$\beta=0.1$. Comparing these plots with
Fig. \ref{lowfreq_eigenfunc}---\ref{lowfreq_eigenfunc_2d}, which have
$\beta=10^{-3}$, we see that increasing $\beta$ restricts the region 
characterized by $W\sim z$ and $\delta v_z\sim$ constant closer to the
mid-plane, with oscillatory behavior emerging near vertical boundaries
where buoyancy becomes important. This leads to the `tilted' pressure
perturbations seen in Fig. \ref{lowfreq_eigenfunc_2d_cool} 
(cf. Fig. \ref{lowfreq_eigenfunc_2d}).

@@

In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable
modes with $\khat=30$. For clarity, we only plot analytic results 
for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the
squares. The fundamental mode is again insensitive boundary
conditions compared to higher order modes, and are in close agreement
with analytic results. As before we observe more effective
stabilization for higher order modes, including the absence of surface
modes (which only occur for $\zmax\geq 5H$) as $\beta$ is increased.  
For $\zmax=3H$, the fundamental mode becomes dominant for $\beta\geq
0.03$. For $\zmax=5H$ and $\zmax=7H$, we
find modes slightly more unstable than the fundamental
mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
panels), but these are clearly associated with boundary
conditions as they are absent for $\zmax=3H$. These observations
suggest the fundamental mode to be more robust in realistic disks.

@@

Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
show that the fundamental mode is not sensitive to boundary
conditions. The fundamental mode also becomes dominant as $\beta$
approaches $\beta_\mathrm{crit}$, except with the possible emergence
of modes associated with boundary conditions. Thus, tracking the
fundamental mode is a more preferable assessment of stability as $\beta$
is increased from zero, rather than searching for the most unstable
mode, because the latter may correspond to modes that depend on boundary
conditions.

>>>>

\subsubsection{Comment on `surface modes'}\label{surf_comment} 
Our fiducial setup with $\khat=10$ in 
Fig. \ref{compare_modes_iso_kx10} (black dots)  lack `surface modes'
found by \cite{nelson13}, \cite{mcnally14}  and 
by \cite{barker15}.  We illustrate surface modes with the mode
diagram for $\khat=30$ in Fig. \ref{compare_modes_iso_kx30}. Surface
modes lie almost parallel to the vertical axis
in such diagrams. They are readily identified for  $\zmax\geq 5H$ and
are marked by `s'.  We typically find an increasing
number of them with increasing $\zmax$ and/or $\khat$,
consistent with \cite{barker15}.

As explained by \cite{barker15}, the occurence of surface modes in the
vertically isothermal disk is due to imposing a finite vertical domain
size. The fastest growing mode is a surface mode, if they
exist, but their association with vertical boundaries means that
$\mathrm{max}(\nu)$ depends on the choice of $\zmax$. Thus, the
vertically isothermal disk is not useful for studying surface modes. 
%Thus, seeking the most unstable mode in a vertically isothermal disk
%may not yield robust results because this

@@

We find the number of unstable modes decrease with increasing
$\beta$. Consider the fiducial setup with $\zmax=5H$ 
(middle panel). The fundamental mode ($M=0$) is not the most unstable for   
$\beta=0.01$ as there are modes with $M>0$ that have larger growth 
rates. However, by comparing with results for $\zmax=3H$, we see that these
higher order modes are strongly influenced by boundary conditions, and
there is a significant mismatch between numerical and analytical
results for $\zmax=3H$ for $M>1$.

However, the fundamental mode becomes dominant for $\beta \geq 
0.03$. Higher order modes, with larger growth rates at small $\beta$,
are more effectively stabilized with increasing $\beta$. For
$\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we only find the
fundamental mode. This is consistent with the discussion in 
\S\ref{iso_vsi_beta_crit}.

@@

Fig. \ref{lowfreq_eigenfunc_cool}---\ref{lowfreq_eigenfunc_2d_cool}
shows the fundamental mode with $\khat=10$ in a disk with 
$\beta=0.1$. Comparing these plots with
Fig. \ref{lowfreq_eigenfunc}---\ref{lowfreq_eigenfunc_2d}, which have
$\beta=10^{-3}$, we see that increasing $\beta$ restricts the region 
characterized by $W\sim z$ and $\delta v_z\sim$ constant closer to the
midplane, with oscillatory behavior emerging near vertical boundaries
where buoyancy becomes important. This leads to the `tilted' pressure
perturbations seen in Fig. \ref{lowfreq_eigenfunc_2d_cool} 
(cf. Fig. \ref{lowfreq_eigenfunc_2d}).

@@

In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable
modes with $\khat=30$. For clarity, we only plot analytic results 
for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the
squares. The fundamental mode is again insensitive boundary
conditions compared to higher order modes, and are in close agreement
with analytic results. As before we observe more effective
stabilization for higher order modes, including the absence of surface
modes (which only occur for $\zmax\geq 5H$) as $\beta$ is increased.  
For $\zmax=3H$, the fundamental mode becomes dominant for $\beta\geq
0.03$. For $\zmax=5H$ and $\zmax=7H$, we
find modes slightly more unstable than the fundamental
mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
panels), but these are clearly associated with boundary
conditions as they are absent for $\zmax=3H$.

@@

Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
show that the fundamental mode is not sensitive to boundary
conditions. The fundamental mode also becomes dominant as $\beta$
approaches $\beta_\mathrm{crit}$, except with the possible emergence
of modes associated with boundary conditions. Thus, tracking the
fundamental mode is a more preferable assessment of stability as $\beta$
is increased from zero, rather than searching for the most unstable
mode. This is because the latter may correspond to modes that depend on boundary
conditions, e.g. surface modes, whose growth rate varies with
$\zmax$.
